Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-22 doc/asplos13/grappa.tex
Further removing implementation from PL view

\paragraph{Global memory access} There are two approaches to {\emph
accessing} global memory. When the programmer expects a computation on
shared data to have spatial locality to exploit, {\em cache} operations
may be used. When there is no locality to exploit, {\em delegate}
operations are used. Since these operations are expected to communicate
with other nodes and have high latency, cache and delegate operations
interact with the scheduler; when issuing a long-latency request, they
suspend the requesting task and allow the core to execute other work
until their response arrives.

\paragraph{Explicit caching} With Grappa's explicit caching support,
applications can instruct Grappa to fetch a global pointer of any length
and return a local pointer to a cached copy of the global memory. Under
the hood, Grappa performs the mechanics of gathering chunks of data from
multiple system nodes and presenting a conventional appearing linear
block of memory as a pointer into a cache. Grappa cache operations have
the usual read-only and read-write variants, along with a write-only
variant used to initialize data structures. Cache operations exploit
spatial locality by reducing the number of small network messages
accessing contiguous data. Languages for distributed shared memory
systems have done optimizations to achieve a similar goal. For example,
the UPC compiler coalesces struct and array accesses into remote get/put
\cite{Chen:2005}, and Fortran D compiler's message vectorization hoists
small messages out of a loop \cite{FortranD:1992}. Caching in Grappa
additionally provides a mechanism for exploiting temporal locality by
operating on the data locally.

\paragraph{Delegate operations}
When the access pattern has low-locality, it is more efficient
to modify the data on its home core rather than bringing a copy to the
requesting core and returning it after modification. Delegate
operations provide this capability. Applications can dispatch
computation to be performed on individual machine-word sized chunks of
global memory to the memory system itself (e.g.,
\emph{fetch-and-add}). \comment{should we bring up delegation in the intro?}

Delegate operations, proposed in \cite{Nelson:hotpar11}, are also the
primary synchronization method in Grappa. Each piece of global memory is
paired with a single core in the system; all accesses to that actual
memory location are done by that core. This allows delegate operations
to be performed atomically on the core simply by making sure no other
code interleaves their operation; and, since context switches in Grappa
are cooperative, this is trivial. Thus, delegate operations are able to
provide atomic semantics to memory owned by one core without using
atomic operations. Using delegation to implement isolation for any
subset of memory has been explored in \cite{delegated:oopsla11}.

>>>>

\paragraph{Global memory access} There are two approaches to {\emph
accessing} global memory. When the programmer expects a computation on
shared data to have spatial locality to exploit, {\em cache} operations
may be used. When there is no locality to exploit, {\em delegate}
operations are used.

\textbf{Explicit caching.} Grappa provides an API to fetch a global
pointer of any length and return a local pointer to a cached copy of the
global memory.  Grappa cache operations have the usual read-only and
read-write variants, along with a write-only variant used to initialize
data structures. Languages for distributed shared memory systems have
done optimizations to achieve a similar goal. For example, the UPC
compiler coalesces struct and array accesses into remote get/put
\cite{Chen:2005}, and Fortran D compiler's message vectorization hoists
small messages out of a loop \cite{FortranD:1992}. Caching in Grappa
additionally provides a mechanism for exploiting temporal locality by
operating on the data locally.

\textbf{Delegate operations.}
When the access pattern has low-locality, it is more efficient to modify
the data on its home core rather than bringing a copy to the requesting
core and returning it after modification. Delegate operations provide
this capability. Applications can dispatch computation to be performed
on individual machine-word sized chunks of global memory to the memory
system itself (e.g., \emph{fetch-and-add}).  Delegate operations,
proposed in \cite{Nelson:hotpar11} and \cite{delegated:oopsla11}, are also the primary synchronization method in Grappa.

### 2012-07-22 doc/asplos13/implementation.tex
Further removing implementation from PL view

@@

\paragraph{Explicit caches} Grappa cache operations are used to exploit
spatial or temporal locality. They allow a single task to buffer a range
of data, read or modify it, and write the modified version back to its
home node(s). All fetch and update operations of cache data are
expressed explicitly by the programmer. Cache operations may operate on
data that spans multiple blocks of the linear address space; the
programmer provides the starting address and length, and the cache
issues as many active messages as necessary to retrieve all necessary
data.

\comment{If anything, this is just a footnote. -M
All current cache operations are incoherent; two tasks may
concurrently modify the same block with their writes interfering
unless protected by synchronization through delegate operations. While
this has been sufficient for our benchmarks, it is possible to embed
synchronization in the cache active messages and reduce the number of
network round trips necessary.}

\subsection{Load Balancing and Network Aggregation}

\subsubsection{Work stealing} \label{subsec:implementation-worksteal}

When a \checkme{grappa-process} runs out of work, it becomes a ``thief''
and asks another other \checkme{grappa-process} (``victim'') for work.
If the victim has tasks to spare in its public queue, they will be
transferred to the thief's public queue.

\comment{
where to put the following related? -BM

Fork join computations: implement with futures or continuation passing
style
feed-forward computations: implement with scalable task counting (we
focus on this style in our benchmarks)
(briefly describe global task joiner)

choices we've made
done -random victims
hmm..  -steal from FIFO side of queue
hmmm... -steal when at the moment readyQ empty and task queues empty
done -steal min(chunk\_size, half)

omit - Recent work has explored scalable work stealing in terms of efficient
termination detection. 
}

\subsubsection{Network Aggregation}

>>>>

\paragraph{Explicit caching.} Under the hood, Grappa performs the
mechanics of gathering chunks of data from multiple system nodes and
presenting a conventional appearing linear block of memory as a local
pointer into a cache. The strategy employed is to issue all the
constituent requests of a cache access request and then yield until all
responses have occurred.  Currently, Grappa caches are \emph{not}
coherent, requiring the programmer to maintain consistent access to
data.  Future work will develop a software directory based coherence
scheme to simplify consistent access to global data.

@@

\subsection{Network Aggregation}
